Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-13 doc/invt.tex
add lj_xerr.gp

As shown in Table 1,
both the inverse-time and optimal schedule reduced
the error significantly.
%
The Gaussian scheme appeared to be superier
to the single-bin scheme in the $\varrho^* = 0.1$ case,
but not in the $\varrho^* = 0.8$ case,
and in the latter case,
an overly large width of the Gaussian
was counterproductive.
%

\begin{table}[h]\footnotesize
  \caption{\label{tab:error_LJ}
    Error of the LJ system.
    The numbers in parentheses are predicted values;
    $\Err_\mathrm{init.}$
    denotes the error
    at the end of the equilibration phase with a constant magnitude;
    $\Err_\mathrm{final}$
    denotes the error at the end of the production phase under
    either the inverse-time ($1/t$) or optimal schedule (opt.).
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Sched. &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    \hline
    \multicolumn{4}{c}{$\varrho^* = 0.1$, $500$ bins} \\
    \hline
    single-bin & $1/t$
    & $5.7(5.7)\times10^{-2}$
    & $9.1(11.4)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $7.7(8.9)\times10^{-3}$
    & $6.8(8.2)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $7.6\times10^{-3}$
    & $3.2\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $6.5\times10^{-3}$
    & $4.0\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $7.0\times10^{-3}$
    & $2.5\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $5.5\times10^{-3}$
    & $3.2\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $5.8\times10^{-3}$
    & $2.6\times10^{-5}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $8.1(9.0)\times10^{-3}$
    & $1.6(1.8)\times10^{-5}$
    \\
    \hline
    \multicolumn{4}{c}{$\varrho^* = 0.8$, $250$ bins} \\
    \hline
    single-bin & $1/t$
    & $1.3(1.4)$
    & $2.7(2.7)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $1.3(1.3)$
    & $2.9(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $1.2$
    & $2.8\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $1.3$
    & $3.4\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $1.2$
    & $3.1\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $0.95$
    & $9.8\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $1.0$
    & $5.8\times10^{-3}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $1.3(1.3)$
    & $3.0(2.7)\times10^{-3}$
    \\
    \hline
  \end{tabular}
\end{table}

@@

\section{\label{sec:Gamma_measure}
Measurement of the integrals of the autocorrelation function
}

@@

It is natural to first introduce homogeneous updating schemes
for a periodic variable\cite{dama2014},
the updating matrix, $\mathbf w$,
assumes the following form
%
\begin{equation}
  w_{ij}
  =
  \mu_{i-j}
  +
  \mu_{i-j+n}
  +
  \mu_{i-j-n}
  ,
\notag
%\label{eq:w_band_pbc}
\end{equation}
%
where the numbers,
$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
characterizing the rigid window
will be referred to as an \emph{updating kernel}\cite{bussi2006}.
%
To satisfy Eq. \eqref{eq:w_sumj},
we impose the normalization
%
\begin{equation}
  \mu_{-b} + \cdots + \mu_b = 1
  ,
\label{eq:mu_normalization}
\end{equation}
%
with $\mu_l = 0$ for $|l| > b$.
%
If $b = n/2$, $\mu_{-b}$ is removed
from the sum in Eq. \eqref{eq:mu_normalization}
to avoid double counting.
%
For simplicity, we also impose the symmetry
%
\begin{equation}
  \mu_i = \mu_{-i}
  .
\label{eq:mu_symm}
\end{equation}

@@

\subsection{Examples}

\subsubsection{Gaussian updating scheme}

@@

\subsubsection{\label{sec:homo_bandpass}
Bandpass updating schemes}

>>>>

As shown in Table 1,
both the inverse-time and optimal schedule reduced
the error significantly,
and the errors roughly matched the predicted values
calculated from the precomputed $\Gamma_k$'s and $\epsilon_k$'s.
%
The Gaussian scheme appeared to be superior
to the single-bin scheme in the $\varrho^* = 0.1$ case,
but not in the $\varrho^* = 0.8$ case.
%
In either case, there was an optimal width of Gaussian
(which was $\sigma = 0.37$ for $\varrho^* = 0.1$
or $0.07$ for $\varrho^* = 0.8$),
an overly large width was counterproductive.
%
In Figure \ref{fig:lj_xerr},
we plotted the error components against the eigenmode index
for the Gaussian updating scheme of width $\sigma = 0.2$.
%
While the Gaussian updating scheme effectively suppressed
the initial errors for short-range (large $k$) modes,
it also slowed down the reduction of errors of these modes.
%
As a result, the final errors of the long-range (small $k$) modes
were larger than those from the single-bin scheme.
%
For the Gaussian updating scheme
the inverse-time schedule tended to reduce the updating magnitude too rapid
leaving a hump on the error profile for the middle-range modes.
%
The optimal schedule improved over the inverse-time schedule
and produced a flatter profile for the final error.

\begin{table}[h]\footnotesize
  \caption{\label{tab:error_LJ}
    Error of the LJ system.
    The numbers have been averaged over multiple runs,
    and the numbers in parentheses are predicted values;
    $\Err_\mathrm{init.}$
    denotes the error
    at the end of the equilibration phase with a constant magnitude;
    $\Err_\mathrm{final}$
    denotes the error at the end of the production phase under
    either the inverse-time ($1/t$) or optimal schedule (opt.).
    \note{The reference values can be found on the second line of \texttt{verr.log}.}
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Sched. &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    \hline
    \multicolumn{4}{c}{$\varrho^* = 0.1$, $500$ bins} \\
    \hline
    single-bin & $1/t$
    & $5.7(5.7)\times10^{-2}$
    & $9.1(11.4)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $7.7(8.9)\times10^{-3}$
    & $6.8(8.2)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $7.6(8.9)\times10^{-3}$
    & $3.2(3.9)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $6.5(7.8)\times10^{-3}$
    & $4.0(4.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $7.0(7.8)\times10^{-3}$
    & $2.5(2.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $5.5(6.6)\times10^{-3}$
    & $3.2(3.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $5.8(6.6)\times10^{-3}$
    & $2.6(3.0)\times10^{-5}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $8.1(8.9)\times10^{-3}$
    & $1.6(1.8)\times10^{-5}$
    \\
    \hline
    \multicolumn{4}{c}{$\varrho^* = 0.8$, $250$ bins} \\
    \hline
    single-bin & $1/t$
    & $1.3(1.4)$
    & $2.7(2.7)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $1.3(1.3)$
    & $2.9(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $1.2(1.3)$
    & $2.8(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $1.3(1.2)$
    & $3.4(3.2)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $1.2(1.2)$
    & $3.1(2.9)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $0.9(0.9)$
    & $9.8(8.0)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $1.0(0.9)$
    & $5.8(5.4)\times10^{-3}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $1.3(1.3)$
    & $3.0(2.7)\times10^{-3}$
    \\
    \hline
  \end{tabular}
\end{table}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_xerr.pdf}
  }
  \caption{
    \label{fig:lj_xerr}
    Error components for the FDS simulation
    using the Gaussian updating scheme with $\sigma = 0.2$.
    The optimal schedule (opt.) is more effective
    in reducing the errors of middle-range modes
    than the inverse-time ($1/t$) schedule.
    The points are joined as a guide to the eyes.
  }
\end{center}
\end{figure}

@@

\section{\label{sec:Gamma_measure}
Measurement of the integrals of the autocorrelation functions of the eigenmodes
}

@@

For a periodic variable\cite{dama2014},
the updating matrix, $\mathbf w$,
assumes the following form
%
\begin{equation}
  w_{ij}
  =
  \mu_{i-j}
  +
  \mu_{i-j+n}
  +
  \mu_{i-j-n}
  ,
\notag
%\label{eq:w_band_pbc}
\end{equation}
%
where the numbers,
$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
characterizing the rigid window
will be referred to as an \emph{updating kernel}\cite{bussi2006}.
%
To satisfy Eq. \eqref{eq:w_sumj},
we impose the normalization
%
\begin{equation}
  \mu_{-b} + \cdots + \mu_b = 1
  ,
\label{eq:mu_normalization}
\end{equation}
%
with $\mu_l = 0$ for $|l| > b$.
%
If $b = n/2$, $\mu_{-b}$ is removed
from the sum in Eq. \eqref{eq:mu_normalization}
to avoid double counting.
%
For simplicity, we also impose the symmetry
%
\begin{equation}
  \mu_i = \mu_{-i}
  .
\label{eq:mu_symm}
\end{equation}

@@

\subsection{Gaussian updating scheme}

@@

\subsection{\label{sec:homo_bandpass}
Bandpass updating schemes}
